|

### 2021-12-14 08-Discussion.Rmd
End of day:

The central aim of this thesis was to infer the causal effect of blood lipid levels on dementia outcomes via evidence synthesis methods. In this discussion, I will review the contribution of each chapter to this aim, summarise the principal findings and methodological innovations of the thesis, and discuss the implications of my findings for both clinical and public health practice. I consider the overall strengths and limitations of this thesis as a coherent body of work and suggest avenues for impactful future research that builds on the research presented.

@@

In __Chapter \@ref(sys-rev-tools-heading)__, I  presented a new tool for the systematic searching of health-related preprints, and detailed two use cases for preprint metadata beyond the systematic searching of this evidence source.

@@

In __Chapter \@ref(cprd-analysis-heading)__, I described an analysis of the association of lipid-regulating agent use and dementia outcomes in the CPRD, a large electronic health record database. I make use of positive and negative control outcomes to illustrate insufficiently controlled confounding.

@@

Overall, I did not identify a consistent effect of any blood lipid fraction (total cholesterol, LDL cholesterol, HDL cholesterol or triglycerides) on any dementia outcome (all-cause dementia, Alzheimer's disease or vascular dementia). While published non-randomised studies provided evidence for a protective effect of lipid-lowering via statin use on Alzheimer's disease and the IPD analysis suggested a harmful effect of raised triglycerides on vascular dementia, these effects were not maintained when incorporated along with other sources of evidence in a quantitative triangulation framework. Similarly, the IPD analysis did not provide evidence for an interaction of patient characteristics with the association of blood lipids on dementia outcomes, though the results are limited by low number of cohorts analysed.

@@

The first is a large-scale analysis of 953,635 patients in the CPRD, followed from their first LDL-c measurement.[@iwagami2021] The study found evidence for a slight increased risk of dementia with higher LDL-c levels at mid-life, driven by the finding for the Alzheimer's disease subgroup. These results are consistent with the findings presented in this thesis, particularly as this study did not have access to _Apo_$\mathcal{E}4$ status and so could not adjust for confounding by this variable. _Apo_$\mathcal{E}4$ is a key risk factor for both Alzheimer's disease and raised LDL-c levels, <!-- TODO CITATION NEEDED --> and so could explain the hamrful association observed. This is supported by a sensitivity analysis reported within this study, which used published information to estimate _Apo_$\mathcal{E}4$ would induce a RR of 1.09 per 1-SD increase in LDL-c in the absence of a true causal effect.

The second analysis examined 502,226 participants in the UK Biobank, a large population-based prospective cohort study.[@gong2021; @sudlow2015] The analysis found no association between lipids measured at mid-life and dementia outcomes, comparable to the results of this thesis. However, similar to the study discussed above, this analysis did not have access to genetic data and so may be subject to residual confounding by _Apo_$\mathcal{E}4$ genotype.

@@

Given this ambiguity, the use of statins should be restricted to their primary and well-understood purpose of lipid-lowering to reduce the risk of cardiovascular events. However, it is important to recognise that even under the scenario that a strong harmful effect of lipid-lowering on dementia outcomes was identified, whether or not to prescribe a statin at mid-life will always be guided by factors other than dementia prevention. Patients must live to a certain age in order to be at risk of dementia (with the exception of familial early-onset dementias), which is substantially less likely if hypercholesterolemia at mid-life is left unchecked.

@@

Similarly, given the ambiguity of evidence, there are no clear implications for public health practice. However, the role of public health, and particularly continuous population surveillance, may be important in providing a further source of evidence in relation to the effect of lipid lowering on dementia. Public health practitioners could examine summary-level statistics of GP prescription data to attempt to relate lipid-lowering treatment to temporal or geographical trends in dementia incidence. It will be difficult to assess if any observed trends are due to a true effect or to confounding or ecological biases. However, as dementia outcomes will be monitored anyway for incidence/prevalence estimation and planning of services/care, this approach would provide a ready source of further evidence with a distinct bias structure to the evidence considered in this thesis.

@@

The proposed quantitative triangulation framework represents the final novel contribution of this thesis to evidence synthesis methodology. Using pre-specified adjustment distributions mapped to the results of the domain-based risk-of-bias and indirectness assessments, this approach provides a systematic way to account for biases/indirectness across studies, reducing the potential for differential adjustment based on knowledge of the results. This framework explcitly seeks to include both direct and indirect evidence, and separates the assessment of bias/indirectness in each result from the definition of distribtuions representing the impact of bias/indirectness.

@@

There are several strengths to this thesis as a whole. Specifically, the identification and triangulation of evidence across study designs and publication status sets this thesis apart from previous synthesis of the evidence on this topic. Additionally, this thesis has also produced new evidence on the association of blood lipids with dementia outcomes in previously unanalysed cohorts and provided additional information on a previously under-studied association (lipids/LRA and vascular dementia).

A further strength of this thesis is the production of software to support the novel evidence synthesis techniques used. Extensive documentation has been written to guide users through usage of the tools and example data is provided for use. Finally, substantial effort has been invested to make the research documented in this thesis as reproducible as possible - this is documented more fully in Appendix \@ref(disc-repro-research).

@@

However, there are also several strong limitations to this thesis. In the first instance, as with many studies of dementia outcomes, a limiting factor in the interpretation of the results is the absence of a detailed pathological mechanism. It is possible that statins have a true effect via some pathway other than lipid-regulation, but without knowledge of the mechanism, the plausibility of this relationship cannot be assessed.

Secondly, each analysis presented here makes use of secondary data sources, be it published literature, electronic health records, or existing cohorts. Secondary data can limit the type of analyses performed (for example, in the IPD analysis presented in Chapter \@ref(ipd-heading), absence of time-to-event data prevented the use of hazard ratios to quantify dementia risk) and the validity of the data if collected for purposes other than research (for example, the accuracy of dementia diagnoses in electronic health records is known to be variable).[@wilkinson2018]

Thirdly, missing evidence was a common limitation across this thesis. In Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading), several studies stated that they did not report the results of an analysis because the results were not significant. This was particularly common among non-randomised studies of exposure examining blood lipids directly. Similarly, an absence of evidence on vascular dementia, potentially due to a publication bias mechanism, limited the analysis of this outcome. Missing evidence was also as an issue in the IPD meta-analysis reported in Chapter \@ref(ipd-heading), in the form of a poor response to data access requests, while the absence of empirically-based distributions of bias/indirectness for use with the triangulation framework limited the credibility of the results produced.

A final weakness stems from the geographical focus of the data analysed in this thesis. All primary analysis presented drew on data from the UK (CPRD, CaPS, EPIC Norfolk, Whitehall II), while the majority of studies identified by the review were based in the Western world (Figure \@ref(fig:cohortLocations)). This may limit the generalisability of the results presented to different populations.

@@

The proposal for the program of research underpinning this thesis was created in advance of starting my PhD, as a detailed plan was required in order to secure the funding that supported me through my studies. In hindsight, the decision to take a broad approach to the inclusion of dementia outcomes resulted in a larger workload than anticipated, particularly when also considering evidence from multiple different study designs. Additionally, as dementia subtypes likely have very different aetiological pathways, it may have been better to focus on a single subtype (e.g. Alzheimer's disease) when considering the causal effects of lipids. Similarly, attempting to undertake a full individual participant data meta-analysis as a single part of a larger program of research was overly ambitious, as data cleaning and harmonisation for just three cohorts was a substantial undertaking.

In the future, armed with the experience gained through conduct of the research projects presented here, I will be better place to scope and design research projects.

@@

Several avenues of future research could be pursued, building on the novel work presented here. These are grouped by topic in the following sections.

@@

It is customary at this point in the discussion of results based (primarily) on observational data to recommend that a large-scale randomised controlled trial be performed to assess the effect of the proposed intervention. Given the costs and logistical challenges associated with trials examining outcomes with long prodomal periods such as dementia, a more efficient approach is offered by the opportunistic post-trial follow-up of existing RCTs of lipid-regulating agents.

@@

Taken together, these points reinforce the need for large-scale GWAS of vascular dementia, similar to those current available for Alzheimer's disease, to identify associated loci that future two-sample Mendelian randomisation studies could exploit. While GWAS of this outcome are likely to be methodologically challenging, given the difficult in diagnosing "pure" vascular dementia, it would be a worthwhile endeavour and allow for the assessment of genetically-driven risk factors beyond those considered in this thesis.

@@

The `medrxivr` tool enables programmatic searching of health-related preprints, and future work could incorporate the tool, along with software to search other literature sources, into an automated searching pipeline as part of a living systematic review.[@elliott2014] In addition, the ready access to preprint data afforded by the tool will enable future meta-epidemiological studies to examine factors which may influence eventual publication (e.g. significant result, geographical location/gender/career stage of first author, or some other factor). Finally, the tool will allow future research to assess the impact of peer review and editorial guidelines on manuscripts through comparison of the same paper at two stages (preprinted versus published). This approach was used to explore the impact of editorial polices on open data sharing, as discussed in Chapter \@ref(sys-rev-tools-heading), but could be applied to other aspects of the publication process.

@@

As noted in Chapter \@ref(sys-rev-results-heading), methods for the inclusion of Mendelian randomisation studies in evidence synthesis exercises are not yet sufficiently developed, particularly in comparison to other study designs. Future work could aim to validate search strategies for this study design, paying particular attention to the range of terms used to define the analytical approach (e.g. genetic instrumental variable analysis). In addition, as discussed in Section \@ref(tri-accuracy), the lack of an established risk-of-bias tool for Mendelian randomisation studies limits subsequent analyses that requires systematic assessment of bias, such as the quantitative triangulation framework proposed here. Creation of a domain-based tool for this study design, which accounts for the differences in potential biases between one- and two-sample approaches, should be considered a priority.

@@

Quantitative triangulation is an area ripe for future work. The approach presented here will benefit from future meta-epidemiological studies to better define the impact of bias/indirectness, as discussed in detail in Section \@ref(tri-prior-def). Similarly, future work could expand the framework proposed here to account for meta-biases, such as those introduced by missing evidence. Tools such as ROB-ME will enable this, but how to adjust for this synthesis-level - as opposed to study-level - bias will need to be determined. Finally, the application of the framework to other causal questions will help to identify sticking points and may lead to refinement of the process.

>>>>

The central aim of this thesis was to infer the causal effect of blood lipid levels on dementia outcomes via evidence synthesis methods. In this discussion, I will review the contribution of each chapter to this aim, summarise the principal findings and methodological innovations of the thesis, and discuss the implications of my findings for both clinical and public health practice. I consider the overall strengths and limitations of this thesis as a coherent body of work and suggest avenues for impactful future research that would build on the research presented.

@@

In __Chapter \@ref(sys-rev-tools-heading)__, I  presented a new tool for the systematic searching of health-related preprints and detailed two use cases for preprint metadata beyond the systematic searching of this evidence source.

@@

In __Chapter \@ref(cprd-analysis-heading)__, I described an analysis of the association of lipid-regulating agent use and dementia outcomes in the CPRD, a large electronic health record database. I made use of positive and negative control outcomes to illustrate insufficiently controlled confounding.

@@

Overall, I did not identify a consistent effect of any blood lipid fraction (total cholesterol, LDL cholesterol, HDL cholesterol or triglycerides) on any dementia outcome (all-cause dementia, Alzheimer's disease or vascular dementia). While published non-randomised studies provided evidence for a protective effect of lipid-lowering via statin use on Alzheimer's disease and the IPD analysis suggested a harmful effect of raised triglycerides on vascular dementia, these effects were not maintained when incorporated along with other sources of evidence in a quantitative triangulation framework. Similarly, the IPD analysis did not provide evidence for an interaction of patient characteristics (age/sex) with the association of blood lipids on dementia outcomes, other than a slight protective effect of male gender on the LDL-c/all-cause dementia relationship. However, this analysis was limited by the low number of cohorts analysed.

@@

The first is a large-scale analysis of 953,635 patients in the CPRD, followed from their first LDL-c measurement.[@iwagami2021] The study found evidence for a slightly increased risk of dementia with higher LDL-c levels at mid-life, driven by the finding for the Alzheimer's disease subgroup. These results are consistent with the findings presented in this thesis, particularly as this study did not have access to _Apo_$\mathcal{E}4$ status and so could not adjust for confounding by this variable. _Apo_$\mathcal{E}4$ is a key risk factor for both Alzheimer's disease and raised LDL-c levels, <!-- TODO CITATION NEEDED --> and so could explain the harmful association observed. This is supported by a simulation analysis reported within this study, which used published information to predict that _Apo_$\mathcal{E}4$ would induce a RR of 1.09 per 1-SD increase in LDL-c in the absence of a true causal effect.

The second analysis examined 502,226 participants in the UK Biobank, a large population-based prospective cohort study.[@gong2021; @sudlow2015] The analysis found no association between lipids measured at mid-life and dementia outcomes, again a comparable finding to the results of this thesis. However, similar to the study discussed above, this analysis did not have access to genetic data and so may be subject to residual confounding by _Apo_$\mathcal{E}4$ genotype.

@@

Given this ambiguity, the use of statins should be restricted to their primary and well-understood purpose of lipid-lowering to reduce the risk of cardiovascular events. However, it is important to recognise that even under a scenario where a strong harmful effect of lipid-lowering on dementia outcomes was identified, whether or not to prescribe a statin at mid-life will always be guided by factors other than dementia prevention. Patients must live to a certain age in order to be at risk of dementia (with the exception of familial early-onset dementias), which is substantially less likely if hypercholesterolemia at mid-life is left unchecked.

@@

Similarly, given the ambiguity of evidence, there are no clear implications for public health practice. However, the role of public health, and particularly continuous population surveillance, may be important in providing a further source of evidence in relation to the effect of lipid lowering on dementia. As an illustrative example, the Cognitive Function and Ageing Study (CFAS) II (2008-2011) showed a reduction in the age-specific prevalence rates of dementia when compared to CFAS I (1989-1994).[@matthews2013] The authors interpreted this reduction as the result of improved education and better control of vascular risk factors, in line with broader thinking on the impact of population-level changes in risk factors.[@capewell2011] In a similar manner, public health practitioners could examine summary-level statistics of GP prescription data to attempt to relate lipid-lowering treatment to temporal or geographical trends in dementia incidence.

It may well prove difficult to assess if any observed trends are due to a true effect or to confounding or ecological biases. However, because dementia outcomes will be monitored anyway for incidence/prevalence estimation and planning of services/care, this approach could provide a ready source of further evidence with a distinct bias structure to the approaches considered in this thesis.

@@

The proposed quantitative triangulation framework represents the final novel contribution of this thesis to evidence synthesis methodology. Using pre-specified adjustment distributions mapped to the results of the domain-based risk-of-bias and indirectness assessments, this approach provides a systematic way to account for biases/indirectness across studies, reducing the potential for differential adjustment based on knowledge of the results. This framework explicitly seeks to include both direct and indirect evidence, and separates the assessment of bias/indirectness in each result from the definition of distributions representing the impact of bias/indirectness.

@@

There are several strengths to this thesis as a whole. Specifically, the identification and triangulation of evidence across study designs and publication status sets this thesis apart from previous syntheses of the evidence on this topic. Additionally, this thesis has also produced new evidence on the association of blood lipids with dementia outcomes in previously unanalysed cohorts, and provided additional information on a previously under-studied relationship (lipids/LRA and vascular dementia).

A further strength of this thesis is the production of software to support the novel evidence synthesis techniques used. Extensive documentation has been written to guide users through usage of the tools and example data are provided, enabling future researchers to readily apply the methods described. Finally, substantial effort has been invested to make the research documented in this thesis as reproducible as possible - this is documented more fully in Appendix \@ref(disc-repro-research).

@@

However, there are also several limitations to this thesis. In the first instance, as with many studies of dementia outcomes, a limiting factor in the interpretation of the results is the absence of a detailed pathological mechanism. It is possible that statins have a true effect via some pathway other than lipid-regulation, but without knowledge of the mechanism, the plausibility of this relationship cannot be assessed.

Secondly, each analysis presented here makes use of secondary data sources, be it published literature, electronic health records, or existing cohorts. Secondary data can limit the type of analyses performed (for example, in the IPD analysis presented in Chapter \@ref(ipd-heading), absence of time-to-event data prevented the use of hazard ratios to quantify dementia risk) and the validity of the data if collected for purposes other than research. For example, the accuracy of dementia diagnoses in electronic health records is known to be variable.[@wilkinson2018]

Thirdly, missing evidence was a common limitation across this thesis. In Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading), several studies stated that they did not report the results of an analysis because the results were not significant. This was particularly common among non-randomised studies of exposure examining blood lipids directly. Similarly, an absence of evidence on vascular dementia, potentially due to a publication bias mechanism, limited the analysis of this outcome. Missing evidence was also an issue in the IPD meta-analysis reported in Chapter \@ref(ipd-heading), in the form of a poor response to data access requests. In addition, the absence of empirically based distributions of bias/indirectness for use with the triangulation framework limited the credibility of the results produced.

A final weakness stems from the geographical focus of the data analysed in this thesis. All primary analysis presented drew on data from the UK (CPRD, CaPS, EPIC Norfolk, Whitehall II), while the majority of studies identified by the review were based in the Western world (Figure \@ref(fig:cohortLocations)). This, combined with the limited evidence for different ethnicities available even within high income settings, may limit the generalisability of the results presented to different populations.

@@

The proposal for the program of research underpinning this thesis was created in advance of starting my PhD, as a detailed plan was required in order to secure the funding that supported me through my studies. In hindsight, the decision to take a broad approach to the inclusion of dementia outcomes resulted in a larger workload than anticipated, particularly when also considering evidence from multiple different study designs. Additionally, as dementia subtypes likely have very different aetiological pathways, it may have been better to focus on a single subtype (e.g. Alzheimer's disease) when considering the causal effects of lipids. Similarly, attempting to undertake a full individual participant data meta-analysis as a single part of a larger program of research was overly ambitious, as data cleaning and harmonisation for just three cohorts turned out to be a substantial undertaking.

In the future, armed with the experience gained through conducting the research presented here, I will be better placed to scope and design research projects.

@@

Several avenues of future research are proposed, building on the novel work presented here. These are grouped by topic in the following sections.

@@

It is customary at this point in the discussion of results based (primarily) on observational data to recommend that a large-scale randomised controlled trial be performed to assess the effect of the proposed intervention. Given the costs and logistical challenges associated with trials examining outcomes with long prodromal periods such as dementia, a more efficient approach is offered by the opportunistic post-trial follow-up of existing RCTs of lipid-regulating agents.

@@

Taken together, these points reinforce the need for a large-scale GWAS of vascular dementia, similar to those currently available for Alzheimer's disease, to identify associated loci that future two-sample Mendelian randomisation studies could exploit. While GWAS of this outcome are likely to be methodologically challenging, given the difficulty in diagnosing "pure" vascular dementia, it would be a worthwhile endeavour and allow for the assessment of genetically driven risk factors beyond those considered in this thesis.

@@

The `medrxivr` tool enables programmatic searching of health-related preprints, and future work could incorporate the tool, along with software to search other literature sources, into an automated searching pipeline as part of a living systematic review.[@elliott2014] In addition, the ready access to preprint data afforded by the tool will enable future meta-epidemiological studies to examine factors which may influence eventual publication (for example,  significant result, geographical location/gender/career stage of first author, or some other factor). Finally, the tool will allow future research to assess the impact of peer review and editorial guidelines on manuscripts through comparison of the same paper at two stages (preprinted versus published). This approach was used to explore the impact of editorial polices on open data sharing, as discussed in Chapter \@ref(sys-rev-tools-heading), but could be applied to other aspects of the publication process.

@@

As noted in Chapter \@ref(sys-rev-results-heading), methods for the inclusion of Mendelian randomisation studies in evidence synthesis exercises are not yet sufficiently developed, particularly in comparison to other study designs. Future work could aim to validate search strategies for this study design, paying particular attention to the range of terms used to define the analytical approach (e.g. genetic instrumental variable analysis). In addition, as discussed in Section \@ref(tri-accuracy), the lack of an established risk-of-bias tool for Mendelian randomisation studies limits subsequent analyses that require systematic assessment of bias, such as the quantitative triangulation framework proposed here. Creation of a domain-based tool for this study design, which accounts for the differences in potential biases between one- and two-sample approaches, should be considered a priority.

@@

Quantitative triangulation is an area ripe for future work. The approach presented here will benefit from future meta-epidemiological studies to better define the impact of bias/indirectness, as discussed in detail in Section \@ref(tri-prior-def). Similarly, future work could expand the framework proposed here to account for meta-biases, such as those introduced by missing evidence. Tools such as ROB-ME will enable this, but the optimum method to adjust for this synthesis-level - as opposed to study-level - bias will need to be determined. Finally, the application of the framework to other causal questions will help to identify sticking points and may lead to refinement of the process.

### 2021-12-14 front-and-back-matter/99-appendices.Rmd
End of day:

`r if(knitr:::is_latex_output()) '\\startappendices'`

@@

Enforcing a minimum period of follow-up is normally used to reduce the risk of reverse causation in an analysis by excluding events occurring within a certain time-frame post-exposure. With respect to dementia outcomes, which have a long prodomal period, enforcing a minimum period of follow-up can add biological validity to an analysis. Incident cases of dementia that occur within a few years of starting a drug may already be beyond effective preventive treatment at the time the drug was started, and so restricting the analysis to those with longer exposure to treatment may provide a more reliable estimate of the effect.

Panel A of Figure \@ref(fig:minFU) describes a standard analysis of an exposure with a minimum enforced follow-up period (in this case, 10 years). Participants enter the cohort and are classified at that point as either exposed or unexposed, for example, on the basis of a total cholesterol measurement. Participants are followed-up over time, but events occuring before the minimum period of follow-up are discarded.

>>>>

---
output:
  word_document: default
  html_document: default
---
`r if(knitr:::is_latex_output()) '\\startappendices'`

@@

Enforcing a minimum period of follow-up is normally used to reduce the risk of reverse causation in an analysis by excluding events occurring within a certain time-frame post-exposure. With respect to dementia outcomes, which have a long prodromal period, enforcing a minimum period of follow-up can add biological validity to an analysis. Incident cases of dementia that occur within a few years of starting a drug may already be beyond effective preventive treatment at the time the drug was started, and so restricting the analysis to those with longer exposure to treatment may provide a more reliable estimate of the effect.

Panel A of Figure \@ref(fig:minFU) describes a standard analysis of an exposure with a minimum enforced follow-up period (in this case, 10 years). Participants enter the cohort and are classified at that point as either exposed or unexposed, for example, on the basis of a total cholesterol measurement. Participants are followed-up over time, but events occurring before the minimum period of follow-up are discarded.

### 2021-12-14 front-and-back-matter/_abstract.Rmd
End of day:

__Results:__ The systematic review did not identify a consistent effect of blood lipids on any dementia outcome across study designs, through there was some suggestion of a protective effect of LDL-c lowering on all-cause dementia and Alzheimer’s disease in observational studies of statin use. The analysis of lipid regulating agents in the CPRD provided weak evidence for a protective effect of statins on all-cause dementia and Alzheimer's disease but suggested a harmful association with vascular dementia. However, the use of control outcomes illustrated this finding was likely due to confounding by indication related to vascular factors. The individual participant data meta-analysis suffered from a low response rate to data access requests, but in the three cohorts analysed, there was some evidence only for the association of triglycerides and vascular dementia. Finally, the triangulation analysis integrating the results of the previous three studies did not provide strong evidence for the causal effect of blood lipids on dementia outcomes.

>>>>

__Results:__ The systematic review did not identify a consistent effect of blood lipids on any dementia outcome across study designs, through there was some suggestion of a protective effect of LDL-c lowering on all-cause dementia and Alzheimer's disease in observational studies of statin use. The analysis of lipid regulating agents in the CPRD provided weak evidence for a protective effect of statins on all-cause dementia and Alzheimer's disease but suggested a harmful association with vascular dementia. However, the use of control outcomes illustrated this finding was likely due to confounding by indication related to vascular factors. The individual participant data meta-analysis suffered from a low response rate to data access requests, but in the three cohorts analysed, there was some evidence only for the association of triglycerides and vascular dementia. Finally, the triangulation analysis integrating the results of the previous three studies did not provide strong evidence for the causal effect of blood lipids on dementia outcomes.

### 2021-12-14 front-and-back-matter/_acknowledgements.Rmd
End of day:

<!-- TODO CHECK ALL NAME SPELLINGS -->

@@

@@

Thanks must also go to those who contributed their time and energy to my research projects: Matthew Lee, Lena Schmidt, Alexandra McAleenan, and Vincent Cheng. Francesca Spiga kindly shared an advance copy of her review of tools to assess Mendelian randomisation studies, and Rebecca Turner shared her code for bias-/indirectness-adjusted meta-analyses that underpins the quantitative triangulation framework presented in Chapter \@ref(tri-heading). Thanks to Vivienne for proof-reading.

@@

On a personal note, there are several people without whom I would not have survived the process of writing a dissertation during COVID times. Thanks to Mum and Dad for your never-ending love and support (I promise I'll get a real job now!). Thank also to Doug for introducing me to the joys of running, Ben for always being around to chat, and Christin for encouraging regular coffee-breaks.

@@

&nbsp;
&nbsp;
&nbsp;
&nbsp;
&nbsp;
&nbsp;
&nbsp;
&nbsp;

>>>>

<!-- TODO UPDATE and check all name spellings -->

__Professional__

@@

Thank Julian separately -

@@

Thanks must also go to those who contributed their time and energy to my research projects: Matthew Lee, Lena Schmidt, Alexandra McAleenan, and Vincent Cheng. Francesca Spiga kindly shared an advance copy of her review of tools to assess Mendelian randomisation studies, and Rebecca Turner shared her code for bias-/indirectness-adjusted meta-analyses that underpins the quantitative triangulation framework presented in Chapter \@ref(tri-heading).

@@

\bigskip

__Personal__

This thesis would not have been possible with Ciara Gardiner. Lots of people say that at this point in the Acknowledgements, but I mean it literally. Between here managing.

The person I have loved with all my heart

Without here

@@

Thanks to Vivienne and Dad for giving up a substantial amount of time to proof-reading.

On a personal note, there are several people without whom I would not have survived the process of writing a dissertation during COVID times.

Thank also to Doug for introducing me to the joys of running, Ben for always being around to chat, and Christin for encouraging regular coffee-breaks.

Thanks to Mum and Dad for your never-ending love and support, and everything they have done to get me to point there doing and submitting a thesis.

\bigskip

### 2021-12-14 index.Rmd
End of day:

author: Luke A McGuinness
college: 
degree: Doctor of Philosophy
degreedate: December 2021
word-count: |
  `r read.csv("data/words/words.csv") %>% dplyr::pull(1) %>% dplyr::last() %>% comma()`
abstract: |
  `r paste(readLines("front-and-back-matter/_abstract.Rmd"), collapse = '\n  ')`
acknowledgements: |
  `r paste(readLines("front-and-back-matter/_acknowledgements.md"), collapse = '\n  ')`
declaration: |
  `r paste(readLines("front-and-back-matter/_declaration.Rmd"), collapse = '\n  ')`
dedication: For Brendan McHugh
covidStatment: |
  `r paste(readLines("front-and-back-matter/_covidStatment.Rmd"), collapse = '\n  ')`
abbreviations: "front-and-back-matter/abbreviations" # path to .tex file with abbreviations

>>>>

author: Luke A McGuinness
college: 
degree: Doctor of Philosophy
degreedate: December 2021
word-count: |
  `r read.csv("data/words/words.csv") %>% dplyr::pull(1) %>% dplyr::last() %>% comma()`
abstract: |
  `r paste(readLines("front-and-back-matter/_abstract.Rmd"), collapse = '\n  ')`
acknowledgements: |
  `r paste(readLines("front-and-back-matter/_acknowledgements.md"), collapse = '\n  ')`
declaration: |
  `r paste(readLines("front-and-back-matter/_declaration.Rmd"), collapse = '\n  ')`
dedication: |
  For\newline
  Brendan McHugh\newline
  &\newline
  Ciara Gardiner
covidStatment: |
  `r paste(readLines("front-and-back-matter/_covidStatment.Rmd"), collapse = '\n  ')`
abbreviations: "front-and-back-matter/abbreviations" # path to .tex file with abbreviations
